Security vs. Maintainability: \\ Fixing Vulnerabilities Obfuscates your Code

### 2019-01-28 paper/ICPC19/icpc19.tex
pass S3

There are security flaws that are more difficult to refactor than others. For
instance, introducing a mechanism to secure authentication is not so easy to
implement as fixing a cross-site scripting vulnerability with a \texttt{escape}
function. Knowing which patterns are more likely to increase maintainability
issues is one step forward to bring awareness to security engineers of what
patterns need more attention. The taxonomy of security patterns used to answer
this question is the same as the one provided by the authors of the dataset.
Maintainability is measured separately for each pattern.
%
\section{Methodology}\label{sec:methodology}
%
The methodology used to measure the impact of security refactorings on the
maintainability of open-source software is discussed in the following sections
and illustrated in Figure \ref{fig:met}. Aiming to answer the research questions
outlined in the previous section, in this paper, we use a dataset containing
$716$ security refactorings collected from open-source software available on
GitHub. $648$ of those refactorings were reported in a previous study conducted
in 2017 \cite{Reis:2017:IJSSE}. Moreover, a baseline of regular commits was
randomly collected from the list of projects of the main dataset to evaluate
what is the impact of regular commits on the maintainability of open-source
software. The Software Improvement Group's\footnote{SIG's website is available
at https://www.sig.eu/ (Accessed on \today{})} web-based source code
analysis service Better Code Hub (BCH)\footnote{BCH's website is available at
https://bettercodehub.com/ (Accessed on \today{})} was used to quantify
maintainability for both security and regular commits.

\begin{figure}[h]
 	\centering 	\includegraphics[width=0.5\textwidth]{figures/methodology.pdf}
 	\caption{Study Methodology}
	\label{fig:met}
\end{figure}

\subsection{Dataset}

To evaluate the impact of security refactorings on the maintainability of open-source software, we use a dataset of security flaws which is the outcome of mining more than $248$ GitHub\footnote{Github's website available at https://github.com/ (Accessed on January 23, 2019)} projects. Reis and Abreu (2017) mined open-source software aiming the extraction of real test cases - created by real developers on their daily basis development - to test and assess the performance of static analysis tools since using hand-seeded test cases or mutations could lead to misleading assessments of the capabilities of the tools. The study yielded to a dataset of $648$ test cases for $16$ different patterns. Each test case of the dataset is a triplet of folders: the commit before the refactoring, the commit responsible for the refactoring and the snippets of code that differ from one version to another (usually, called \textit{diff}) - where one can easily identify the code used to fix the security flaw. In this study, we focus on computing the maintainability of the commits before and after the security refactoring to evaluate if the impact was positive, negative or none.

@@

Reis and Abreu ($2017$) dataset commits were manually validated, flagged as security flaws and classified with one of the $16$ following patterns: \textit{Injection}, \textit{Broken Authentication and Session Management}, \textit{Cross-Site Scripting}, \textit{Broken Access Control}, \textit{Security Misconfiguration}, \textit{Sensitive Data Exposure}, \textit{Insufficient Attack Protection}, \textit{Cross-Site Request Forgery}, \textit{Using Components With Known Vulnerabilities}, \textit{Unprotected APIs}, \textit{Memory Leak}, \textit{Overflow}, \textit{Denial-of-Service}, \textit{Path Traversal} and \textit{Miscellaneous}. However, the dataset evolved and now integrates new patterns, such as \textit{Resource Leak} and \textit{SHA-1 Hash Function}. A considerable amount of these patterns are based in the OWASP Top 10 of 2013~\cite{oswap:2013} and OWASP Top 10 of 2017~\cite{oswap:2017}.

In total, the dataset contains $716$ security refactorings. These refactorings were computed by the BCH tool to calculate maintainability. However, some of the refactorings are not considered in the final results due to limitations of BCH: lack of language support and project size. The final dataset contains $607$ security refactorings. The Wilcoxon statistical test - described more in-depth in subsection \ref{sec:statsval} - is performed to evaluate the significance of our results. Each type of security flaw is tested, Figure \ref{fig:pat}. One of the main requirements of this test is that the input sample needs to comprise more than $20$ commits. Thus, the patterns not complying with this requirement were reallocated to the \textit{Miscellaneous} pattern. In the end, we highlight the following patterns:

\begin{itemize}
	\item \textbf{Using Components with Known Vulnerabilities.} The majority of software produced today integrates several components, such as libraries and frameworks, which may affect the software if developers use vulnerable versions - usually known and disclosed somewhere on the Internet. \textit{Based on the OWASP Top 10 of 2017.}
	\item \textbf{Broken Authentication \& Session Management.} Uncorrectly implemented functionalities related to authentication and session management, allowing crackers to gain access to session tokens, passwords, keys and other sensitive data. \textit{Based on the OWASP Top 10 of 2013.}
	\item \textbf{Cross-Site Request Forgery.} Poor session tokens generation and management usually allow crackers to send forged HTTP requests including authentication information from the victim to vulnerable web application. \textit{Based on the OWASP Top 10 of 2013.}
	\item \textbf{Denial-of-Service} Security flaws that can allow a cracker to flood or crashing services. This attacks usually occur when the system receives much traffic causing it to slow down and eventually stop. For example, flaws allowing crackers to trigger memory allocations corresponding to large length values (Listing \ref{lst:vuln}).
	\item \textbf{Miscellaneous} This pattern comprises several others categories of security refactorings (e.g., \textit{path traversal and buffer overflow}). It contains several other security refactorings that do not have their own pattern yet. The patterns not satisfying the size requirement of more than 20 refactorings are contemplated here.
	\item \textbf{Cross-Site Scripting.} Lack of proper validation or escaping allow crackers to submit untrusted data to web browsers through malicious scripts that can hijack the user sessions or redirect the user to malicious sites. \textit{Based on the OWASP Top 10 of 2017.}
	\item \textbf{Injection.} When developers do not keep untrusted data separate from commands and queries. If a cracker sends a string that exploits the syntax of the interpreter, then an injection attack is possible (e.g., SQL and LDAP injection). \textit{Based on the OWASP Top 10 of 2017.}
	\item \textbf{Memory Leak.} Memory management issue found more frequently in programming languages that do not manage memory automatically (e.g., C/C++ and Objective-C), i.e., where instead developers are responsible for handling it. One of the main causes of DoS attacks.
\end{itemize}

The dataset contemplates security flaws for more than 13 different languages being PHP (39\%) and C (24\%) the most prevalent ones, Figure \ref{fig:lang}.  To classify the software of the dataset, we use a taxonomy for open-source software published in an older study~\cite{7816479}. Figure \ref{fig:domain} presents the projects domain distribution: \textit{Aplication Software} ($25$), software that provides end-users with functional systems; \textit{Web libraries and frameworks} ($20$); \textit{System Software} ($18$), software that provides services and infrustructures (e.g., operating systems, servers and databases); \textit{Software Tool} ($13$), software that supports development (e.g., programming languages, compilers, package managers, IDEs); and, \textit{Non-web libraries and frameworks} ($18$), for desktop and mobile software.

\begin{table*}[h]
	\centering
	\caption{Descriptive statistics of the dataset projects}
\begin{tabular}{@{}lllllllllll@{}}
\toprule
      & forks   & stars   & watchers & contributors & commits  & branches & releases & size      & issues & pull requests  \\ \midrule
mean  & 1763.52 & 5448.74 & 401.37   & 153.33       & 14834.17 & 45.17    & 129.45   & 122973.24 & 3768.97   & 1941.61 \\
std   & 2434.03 & 6215.09 & 486.60   & 123.68       & 22234.46 & 150.15   & 189.93   & 209732.51 & 5933.16   & 3603.31 \\
min   & 1       & 3       & 1        & 0            & 103      & 1        & 0        & 108       & 0         & 0       \\
25\%  & 391.50  & 1581    & 117.25   & 49           & 1440.50  & 4        & 19       & 8466.75   & 313.75    & 143.25  \\
median  & 838.50  & 2836.50 & 248      & 99           & 5504.50  & 9        & 59       & 37372.50  & 1792.50   & 650     \\
75\%  & 2155    & 6828.50 & 459.50   & 261          & 18579.25 & 20       & 142.75   & 117699.50 & 4087.75   & 1907.25 \\
max   & 16366   & 31841   & 3446     & 413          & 114378   & 1227     & 1114     & 995790    & 33970     & 19329   \\
Total & 165771  & 512182  & 37729    & 14413        & 1394412  & 4246     & 12168    & 11559485  & 354283    & 182511  \\ \bottomrule
\end{tabular}
\label{tab:dataset}
\end{table*}

Table \ref{tab:dataset} presents the descriptive statistics of the 94 open-source projects involved in this study. \Sof{@all: Should I expand this? I am not sure what it is important to say about this.}

\subsection{Security vs. Baseline Commits}

Previous studies tried to measure the impact of refactorings on open-source software maintainability~\cite{HEGEDUS2018313} before. But to the best of our knowledge, none of them used the BCH model and tried to evaluate the impact of security refactorings compared with regular refactorings as we do. Thus, we analyze the maintainability of regular commits and use them as a baseline.

The baseline dataset uses the security commits dataset as input. For each
security commit, one random commit is selected from the list of all commits of
the belonging project. We originate the regular commits from the security
commits to ensure that differences in maintainability are not consequence of
characteristics of different projects.

\subsection{Maintainability Analysis}

Better Code Hub is used to collect the maintanability reports of the refactorings of each project. Figure \ref{fig:guidelines} presents the 10 different guidelines proposed by BCH authors for delivering software that is easy to maintain \cite{Visser:2016:OREILLY}. The reports provided by the tool are based on the following guidelines:

@@

All guidelines are measured in the same evaluation by BCH. During each guideline evaluation, the tool determines the compliance towards one guideline by establishing limits for the percentage of code allowed to be in each of the 4 risk severity levels (\emph{low risk}, \emph{medium risk}, \emph{high risk} and \emph{very high risk}). If the project does not violate the thresholds, then it is compliant with the guideline. These thresholds are calibrated by BCH using their own data - open-source and closed software systems. If a project is compliant with a guideline, it means that it is at least $65\%$ better than the software used by BCH to calculate the thresholds\footnote{Check the answer to \emph{How can I adjust the threshold for passing/not passing a guideline?} at https://bettercodehub.com/docs/faq (Accessed on January 27, 2019)}. Figure \ref{fig:bchrep} shows an example of the  guidelines report provided by BCH for a project after finishing its evaluation. These are the results after performing the OpenSSL CVE-2014-3506 vulnerability refactoring - motivation example discussed, previously, in section \ref{sec:motivation}. This version of OpenSSL only complies with 1 out of 10 guidelines, the \emph{Write Clean Code} guideline.

@@

\subsection{Statistical Validation}\label{sec:statsval}

To validate the maintainability differences in different groups of commits (e.g., baseline and security commits), we use the Paired Wilcoxon signed-rank test \cite{10.2307/3001968} with the significance level $\alpha = 0.05$. In other words, we test the null hypothesis that the maintainability difference between pairs of versions $v_{v-1}$, $v_v$ (i.e., before and after a security-commit) follows a symmetric distribution around $0$. However, when the difference between groups is zero, the observations are discarded. Thus, we use a more robust approach of the wilcoxon test, the Wilcoxon test by Pratt which provides an alternative that incorporates the cases where the maintainability difference is zero \cite{10.2307/2282543}. To understand the effect-size, as advocated by the Common-language effect sizes \cite{graw:1992}, we compute the mean difference, the median of the difference, and the percentage of cases that reduce maintainability.

\section{Results}\label{sec:results}

>>>>

There are security flaws that are more difficult to refactor than others. For
instance, introducing a mechanism to secure authentication is not so easy to
implement as fixing a cross-site scripting vulnerability with a \texttt{escape}
function. Knowing which patterns are more likely to increase maintainability
issues is one step forward to bring awareness to security engineers of what
patterns need more attention. The taxonomy of security patterns used to answer
this question is the same as the one provided by the authors of the dataset.
Maintainability is measured separately for each pattern.
%
\section{Methodology}\label{sec:methodology}
%
The methodology used to measure the impact of security refactorings on the
maintainability of open-source software is discussed in the following sections
and illustrated in Figure \ref{fig:met}. Aiming to answer the research questions
outlined in the previous section, in this paper, we use a dataset containing
$716$ security refactorings collected from open-source software available on
GitHub~\cite{Reis:2017:IJSSE}. Moreover, a baseline of regular commits was
randomly collected from the list of projects of the main dataset to evaluate
what is the impact of regular commits on the maintainability of open-source
software. The Software Improvement Group's\footnote{SIG's website is available
at https://www.sig.eu/ (Accessed on \today{})} web-based source code analysis
service \emph{Better Code Hub} (BCH)\footnote{BCH's website is available at
https://bettercodehub.com/ (Accessed on \today{})} was used to quantify
maintainability for both security and regular commits.
%
\begin{figure}[h]
 	\centering 	\includegraphics[width=0.5\textwidth]{figures/methodology.pdf}
 	\caption{Study Methodology}
	\label{fig:met}
\end{figure}
%
\subsection{Dataset}
%
To evaluate the impact of security refactorings on the maintainability of
open-source software, we use a dataset of security flaws which is the outcome of
mining $94$ GitHub projects. Reis and Abreu (2017) mined open-source
software aiming at the extraction of real examples --- created by real
developers --- to test and assess the performance of static analysis tools since
using hand-seeded test cases or mutations could lead to misleading assessments
of the capabilities of the tools \cite{just2014mutants}. The study yielded a
dataset with $716$ test case for $16$ security patterns. Each test case of the
dataset is a triplet: the commit before the refactoring, the commit responsible
for the refactoring and the snippets of code that differ from one version to
another (typically, called \textit{diff}) --- where one can easily identify the
code used to fix the security flaw. In this study, we focus on computing the
maintainability of the commits before and after the security refactoring to
evaluate if the impact was positive, negative or none.

@@

The test cases in the dataset are classified with one of the $16$ following
patterns, as based in the OWASP Top 10 of 2013~\cite{oswap:2013} and OWASP Top
10 of 2017~\cite{oswap:2017}: \textit{Injection}, \textit{Broken Authentication
and Session Management}, \textit{Cross-Site Scripting}, \textit{Broken Access
Control}, \textit{Security Misconfiguration}, \textit{Sensitive Data Exposure},
\textit{Insufficient Attack Protection}, \textit{Cross-Site Request Forgery},
\textit{Using Components With Known Vulnerabilities}, \textit{Unprotected APIs},
\textit{Memory Leak}, \textit{Overflow}, \textit{Denial-of-Service},
\textit{Path Traversal}, \textit{Miscellaneous}, \textit{Resource Leak} and
\textit{SHA-1 Hash Function}.

The $716$ refactorings in the dataset were run against the BCH toolset to
calculate their maintainability reports. However, there are refactorings
discarded from tour study due to limitations of BCH: in particular, lack of
language support and project size. The final dataset used in this paper contains
$607$ security refactorings, labelled as one of the following categories (we
have decided to change the label to \textit{Miscellaneous} of the patterns in
the dataset containing less than $20$ instances):

\begin{itemize}
	\item \textbf{Using Components with Known Vulnerabilities.} The majority of
	software produced today integrates several components, such as libraries and
	frameworks, which may affect the software if developers use vulnerable
	versions - usually known and disclosed somewhere on the Internet.
	\textit{Based on the OWASP Top 10 of 2017.}

	\item \textbf{Broken Authentication \& Session Management.} Uncorrectly
	implemented functionalities related to authentication and session management,
	allowing crackers to gain access to session tokens, passwords, keys and other
	sensitive data. \textit{Based on the OWASP Top 10 of 2013.}

	\item \textbf{Cross-Site Request Forgery.} Poor session tokens generation and
	management usually allow crackers to send forged HTTP requests including
	authentication information from the victim to vulnerable web application.
	\textit{Based on the OWASP Top 10 of 2013.}

	\item \textbf{Denial-of-Service} Security flaws that can allow a cracker to
	flood or crashing services. This attacks usually occur when the system
	receives much traffic causing it to slow down and eventually stop. For
	example, flaws allowing crackers to trigger memory allocations corresponding
	to large length values (Listing \ref{lst:vuln}).

	\item \textbf{Miscellaneous} This pattern comprises several others categories
	of security refactorings (e.g., \textit{path traversal and buffer overflow}).
	It contains several other security refactorings that do not have their own
	pattern yet. The patterns not satisfying the size requirement of more than 20
	refactorings are contemplated here.

	\item \textbf{Cross-Site Scripting.} Lack of proper validation or escaping
	allow crackers to submit untrusted data to web browsers through malicious
	scripts that can hijack the user sessions or redirect the user to malicious
	sites. \textit{Based on the OWASP Top 10 of 2017.}

	\item \textbf{Injection.} When developers do not keep untrusted data separate
	from commands and queries. If a cracker sends a string that exploits the
	syntax of the interpreter, then an injection attack is possible (e.g., SQL and
	LDAP injection). \textit{Based on the OWASP Top 10 of 2017.}

	\item \textbf{Memory Leak.} Memory management issue found more frequently in
	programming languages that do not manage memory automatically (e.g., C/C++ and
	Objective-C), i.e., where instead developers are responsible for handling it.
	One of the main causes of DoS attacks.
\end{itemize}

The dataset contemplates security flaws for more than 13 different languages
being PHP (39\%) and C (24\%) the most prevalent ones; see Figure \ref{fig:lang}.
To classify the applications in the dataset, we use a taxonomy for open-source
software published in an older study~\cite{7816479}. Figure \ref{fig:domain}
presents the projects domain distribution: \textit{Aplication Software} ($25$),
software that provides end-users with functional systems; \textit{Web libraries
and frameworks} ($20$); \textit{System Software} ($18$), software that provides
services and infrustructures (e.g., operating systems, servers and databases);
\textit{Software Tool} ($13$), software that supports development (e.g.,
programming languages, compilers, package managers, IDEs); and, \textit{Non-web
libraries and frameworks} ($18$), for desktop and mobile software.
%
Table \ref{tab:dataset} presents the descriptive statistics of the $94$
open-source projects involved in this study, icluding number of forks, number of
stars, number of watchers, number of contributors, number of commits, number of
branches, number of releases,  size (cf. GitHub it is the size of the whole
repository, including all of its history, in kilobytes), number of issues and pull
requests.

\begin{table*}[h]
	\centering
	\caption{Descriptive statistics of the dataset projects}
\begin{tabular}{@{}lllllllllll@{}}
\toprule
      & forks   & stars   & watchers & contributors & commits  & branches & releases & size      & issues & pull requests  \\ \midrule
mean  & 1763.52 & 5448.74 & 401.37   & 153.33       & 14834.17 & 45.17    & 129.45   & 122973.24 & 3768.97   & 1941.61 \\
std   & 2434.03 & 6215.09 & 486.60   & 123.68       & 22234.46 & 150.15   & 189.93   & 209732.51 & 5933.16   & 3603.31 \\
min   & 1       & 3       & 1        & 0            & 103      & 1        & 0        & 108       & 0         & 0       \\
25\%  & 391.50  & 1581    & 117.25   & 49           & 1440.50  & 4        & 19       & 8466.75   & 313.75    & 143.25  \\
median  & 838.50  & 2836.50 & 248      & 99           & 5504.50  & 9        & 59       & 37372.50  & 1792.50   & 650     \\
75\%  & 2155    & 6828.50 & 459.50   & 261          & 18579.25 & 20       & 142.75   & 117699.50 & 4087.75   & 1907.25 \\
max   & 16366   & 31841   & 3446     & 413          & 114378   & 1227     & 1114     & 995790    & 33970     & 19329   \\
Total & 165771  & 512182  & 37729    & 14413        & 1394412  & 4246     & 12168    & 11559485  & 354283    & 182511  \\ \bottomrule
\end{tabular}
\label{tab:dataset}
\end{table*}
%
\subsection{Security vs. Baseline Commits}
%
Previous studies attempted to measure the impact of code refactorings on
open-source software maintainability~\cite{HEGEDUS2018313} before. But, to the
best of our knowledge, no previous work has used the BCH model and focused on
evaluating the impact of security refactorings and compare with regular
refactorings as in this study. We analyze the maintainability of regular commits
(i.e., commits not related with security fixes) and use them as a baseline.

The baseline dataset uses the security commits dataset as input. For each
security commit, one random commit is selected from the list of all commits of
that particular project. We originate the regular commits from the security
commits to ensure that differences in maintainability are not consequence of
characteristics of different projects.
%
\subsection{Maintainability Analysis}

As saids before, the web-based source code analysis service \emph{Better Code
Hub} (BCH) is used to collect the maintanability reports of the refactorings of
each project. Figure \ref{fig:guidelines} presents the 10 guidelines
proposed by BCH's authors for delivering software that is easy to maintain
\cite{Visser:2016:OREILLY}.

@@

All guidelines are computed in the same evaluation by BCH. During each guideline
evaluation, the tool determines the compliance towards one guideline by
establishing limits for the percentage of code allowed to be in each of the $4$
risk severity levels (\emph{low risk}, \emph{medium risk}, \emph{high risk} and
\emph{very high risk}). If the project does not violate the thresholds, then it
is compliant with the guideline. These thresholds are calibrated by BCH using
their own data/experience --- open-source and closed software systems. If a
project is compliant with a guideline, it means that it is at least $65\%$
better than the software used by BCH to calculate the thresholds\footnote{Check
the answer to \emph{How can I adjust the threshold for passing/not passing a
guideline?} at https://bettercodehub.com/docs/faq (Accessed on \today{})}.
Figure \ref{fig:bchrep} shows an example of the  guidelines report
provided by BCH for a project after finishing its evaluation. These are the
results after performing the OpenSSL CVE-2014-3506 vulnerability refactoring -
motivation example discussed, previously, in section \ref{sec:motivation}. This
version of OpenSSL only complies with 1 out of 10 guidelines, the \emph{Write
Clean Code} guideline.

@@

\subsection{Statistical Validation}\label{sec:statsval}
%
To validate the maintainability differences in different groups of commits
(e.g., baseline and security commits), we use the Paired Wilcoxon signed-rank
test with the significance level $\alpha = 0.05$ \cite{10.2307/3001968} . In
other words, we test the null hypothesis that the maintainability difference
between pairs of versions $v_{v-1}$, $v_v$ (i.e., before and after a
security-commit) follows a symmetric distribution around $0$. However, when the
difference between groups is zero, the observations are discarded. Thus, we use
a more robust approach of the wilcoxon test, the Wilcoxon test by Pratt which
provides an alternative that incorporates the cases where the maintainability
difference is zero \cite{10.2307/2282543}. To understand the effect-size, as
advocated by the Common-language effect sizes \cite{graw:1992}, we compute the
mean difference, the median of the difference, and the percentage of cases that
reduce maintainability.
%
\section{Results}\label{sec:results}
